---
title: "Yield projections_other crops"
author: "Namita Joshi"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true                # Enables the table of contents
    toc_float: true          # Makes the TOC a floating sidebar
    code_folding: hide       # Hides code by default
    includes:
      after_body: hypothes.html  # This points to the custom HTML file
vignette: >
  %\VignetteIndexEntry{Legume yield projections}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
comments:
  utterances:
    repo: Namita-J/climate-impacts-legumes
---

```{r}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

```{r packages,include=F,eval=T,echo=F}

pacman::p_load(
   knitr, readxl, httr, leaflet, dplyr, shiny, ggplot2, tidyr, sf, rnaturalearth,
   rnaturalearthdata, RColorBrewer, ggridges
)
```

```{r reading in the data, include=F,eval=T,echo=T }
# Set a directory for downloaded data
dl_dir <- "downloaded_data"

# Create the directory if it doesn't already exist
if(!dir.exists(dl_dir)){
  dir.create(dl_dir, showWarnings = FALSE)
}

# Define the URL and local path for your file
data_url <- "https://raw.githubusercontent.com/Namita-J/climate-impacts-legumes/main/yield_predictions_data.xlsx"

data_local <- file.path(getwd(), dl_dir, "yield_predictions_data.xlsx")

# Set whether to overwrite the file if it already exists
overwrite <- TRUE

# Check if the file exists and remove it if overwrite is TRUE
if (file.exists(data_local) & overwrite) {
  file.remove(data_local)
}

# Download the file (set 'update=TRUE' to refresh the local copy)
update <- TRUE
if (update) {
  download.file(data_url, data_local, mode = "wb")  # Write in binary mode
}

# Read only the "Dataset" tab from the Excel file
dataset <- readxl::read_excel(data_local, sheet = "Dataset")

# Display a preview of the data
head(dataset)
```

```{r distribution of papers in Africa Static, echo=T, message=FALSE}
excluded_crops <- c("Common bean", "Groundnut", "Cowpea", "Soybean", 
                    "Bush Bean", "Climbing Bean", "Faba Bean", "Green grams", "Chickpea","Common Bean",
                    "Bambara")

# Filter dataset to exclude selected crops
filtered_data <- dataset %>%
  filter(!Crop %in% excluded_crops) %>%
  mutate(
    longitude = as.numeric(longitude),
    latitude = as.numeric(latitude)
  ) %>%
  filter(!is.na(longitude) & !is.na(latitude))  # Ensure valid coordinates

# Count unique DOIs for each crop
crop_study_counts <- filtered_data %>%
  group_by(Crop) %>%
  summarise(Study_Count = n_distinct(doi)) %>%
  ungroup()

# Generate crop labels with study count
filtered_data <- filtered_data %>%
  left_join(crop_study_counts, by = "Crop") %>%
  mutate(Crop_Label = paste0(Crop, " (", Study_Count, ")"))  # Format as "Crop (Count)"

# Convert to spatial data
filtered_data_sf <- st_as_sf(filtered_data, coords = c("longitude", "latitude"), crs = 4326)

# Define a **bold & vibrant** color palette
crop_colors <- c(
  "Bambara" = "red",
  "Cassava" = "blue",
  "Cocoyam" = "limegreen",
  "Common Bean" = "darkorange",
  "Cooking Banana" = "purple",
  "Cotton" = "pink",
  "Maize" = "cyan",
  "Millet" = "gold",
  "Pearl Millet" = "magenta",
  "Pineapple" = "black",
  "Potato" = "darkred",
  "Rice" = "yellow",
  "Sorghum" = "navy",
  "Sugercane" = "coral",
  "Sunflower" = "mediumvioletred",
  "Sweet Banana" = "dodgerblue",
  "Sweet Potato" = "deeppink",
  "Wheat" = "green",
  "Yam" = "turquoise"
)

# Load map data for Africa
africa <- ne_countries(scale = "medium", continent = "Africa", returnclass = "sf")

# Create the map
ggplot() +
  geom_sf(data = africa, fill = "gray95", color = "gray80") +  # Base map
  geom_sf(data = filtered_data_sf, aes(color = Crop), size = 2, alpha = 0.8) +  # Crop points
  scale_color_manual(values = crop_colors) +
  labs(
    title = "Study Locations by Crop",
    color = "Crop"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

# Sort the data from highest to lowest and keep only the top 5
top_crops <- crop_study_counts %>%
  arrange(desc(Study_Count)) %>%
  head(5)

# Display the table using knitr::kable
kable(top_crops, caption = "Top 5 crops with the most papers")
```

```{r box plots of different CROPS, echo=T, message=FALSE}
# Filter dataset to exclude unwanted crops
filtered_data <- dataset %>%
  filter(!Crop %in% excluded_crops) %>%   # Exclude unwanted crops
  filter(!is.na(`Climate impacts (%)`))   # Ensure 'Climate impacts (%)' is not NA

# Count observations for each crop
filtered_data <- filtered_data %>%
  group_by(Crop) %>%
  mutate(Count = n()) %>%
  ungroup()

# Create the boxplot
ggplot(filtered_data, aes(x = `Climate impacts (%)`, y = reorder(Crop, Count, .desc = TRUE))) +
  geom_boxplot(outlier.size = 1, outlier.color = "red") +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "Climate Impact (%)",
    y = "Crop (n)",
    title = "Climate Impact (%) on yield of Selected Crops"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title.y = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    axis.text.y = element_text(size = 10)
  )
```

```{r box plots of different climate scenarios, echo=T, message=FALSE}
# Filter data for specific crops and climate scenarios
filtered_data <- dataset %>%
  filter(!Crop %in% excluded_crops) %>%
  filter(!is.na(`Climate impacts (%)`)) %>%  # Ensure 'Climate impacts (%)' is not NA
  filter(`Climate scenario` %in% c("RCP8.5", "RCP4.5", "RCP2.6", "SSP370", "SSP126", "SSP585"))  # Include only specified scenarios

# Count observations for each climate scenario
filtered_data <- filtered_data %>%
  group_by(`Climate scenario`) %>%
  mutate(Count = n()) %>%
  ungroup()

# Create the boxplot
ggplot(filtered_data, aes(x = `Climate impacts (%)`, y = reorder(`Climate scenario`, Count, .desc = TRUE))) +
  geom_boxplot(outlier.size = 1, outlier.color = "red") +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "Climate Impact (%)",
    y = "Climate Scenario",
    title = "Climate Impact (%) on yield for different Climate Scenarios"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title.y = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    axis.text.y = element_text(size = 10)
  )


```

```{r box plots of different strategies, echo=T, message=FALSE}
# Filter data for specific crops
filtered_data <- dataset %>%
  filter(!Crop %in% excluded_crops) %>%
  filter(!`Adaptation type` %in% "Yield Potential") %>%
  filter(!is.na(`Climate impacts (%)`))  # Ensure 'Climate impacts (%)' is not NA

# Count observations for each crop
filtered_data <- filtered_data %>%
  group_by(`Adaptation type`) %>%
  mutate(Count = n()) %>%
  ungroup()

# Create the boxplot
ggplot(filtered_data, aes(x = `Climate impacts (%)`, y = reorder(`Adaptation type`, Count, .desc = TRUE))) +
  geom_boxplot(outlier.size = 1, outlier.color = "red") +
  geom_vline(xintercept = 0, color = "red", linetype = "dashed") +
  labs(
    x = "Climate Impact (%)",
    y = "Adaptation type",
    title = "Climate Impact (%) on yield with different adaptation strategies"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title.y = element_text(size = 12),
    axis.title.x = element_text(size = 12),
    axis.text.y = element_text(size = 10)
  )
```

The treatment vs. control approach greatly reduces the number of studies included (7), as most focus on a single adaptation strategy rather than projecting both scenarios. This script filters the dataset for selected crops, separating "no adaptation" (control) and specific adaptation strategies (treatment). The two groups are merged based on shared attributes like site, crop, and climate scenario, allowing for direct comparisons.

```{r trt vs control, echo=T, message=FALSE}

# Filter dataset to exclude unwanted crops
filtered_data <- dataset %>%
  filter(!Crop %in% excluded_crops) %>%
  filter(!`Adaptation type` %in% "Yield Potential")

# Create control dataset
control_data <- filtered_data %>%
  filter(`Adaptation type` == "No") %>%
  rename(YieldChange_Control = `Climate impacts (%)`)

# Create treatment dataset
treatment_data <- filtered_data %>%
  filter(`Adaptation type` != "No") %>%
  rename(YieldChange_Treatment = `Climate impacts (%)`)

# Merge control and treatment datasets
merged_data <- merge(
  control_data,
  treatment_data,
  by = c("Ref No", "Future_Mid-point", "Climate scenario",  
         "Site(location)", "Crop","GCM/RCM","Crop Model"),
  all = FALSE
)

# Transform the data for ridge plots
plot_data <- merged_data %>%
  pivot_longer(
    cols = c(YieldChange_Control, YieldChange_Treatment),
    names_to = "Management_Type",
    values_to = "YieldChange"
  ) %>%
  mutate(
    Management_Group = case_when(
      Management_Type == "YieldChange_Control" ~ "No Adaptation",
      TRUE ~ `Adaptation type.y`
    )
  )

# Define **brighter & clearer** color palette
crop_colors <- c(
  "Maize" = "#377EB8",
  "Millet" = "#4DAF4A",
  "Rice" = "#FF7F00",
  "Sorghum" = "#984EA3",
  "Sunflower" = "#FFFF33",
  "Wheat" = "#E41A1C"
)



# Improved Ridge Plot without Dots & with Tick Marks
ggplot(plot_data, aes(
  x = YieldChange,
  y = reorder(Management_Group, YieldChange, median),
  fill = Crop
)) +
  geom_density_ridges(
    scale = 1.5, alpha = 0.7, rel_min_height = 0.01,
    quantile_lines = TRUE, quantiles = 2  # Add median tick marks
  ) +
  scale_fill_manual(values = crop_colors) +  # Use defined colors
  labs(
    title = "Yield Change Distribution by Management Strategy- Trt vs Control",
    x = "Yield Change (%)",
    y = "Management Strategy",
    fill = "Crop"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 12, color = "black"),  # Improve label readability
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.position = "right",
    legend.text = element_text(size = 12)
  )

crop_study_trt_control <- merged_data %>%
  group_by(Crop) %>%
  summarise(Study_Count = n_distinct(doi.x)) %>%
  ungroup()


# Display the table using knitr::kable
kable(crop_study_trt_control, caption = "Number of ppaers used for this plot")

```

**OR A BOX PLOT TO SHOW THE SAME**

```{r BOX PLOT FOR TRT VS CONTROL, echo=T, message=FALSE}
# Filter dataset to exclude unwanted crops
filtered_data <- dataset %>%
  filter(!Crop %in% excluded_crops) %>%
  filter(!`Adaptation type` %in% "Yield Potential")

# Create control dataset
control_data <- filtered_data %>%
  filter(`Adaptation type` == "No") %>%
  rename(YieldChange_Control = `Climate impacts (%)`)

# Create treatment dataset
treatment_data <- filtered_data %>%
  filter(`Adaptation type` != "No") %>%
  rename(YieldChange_Treatment = `Climate impacts (%)`)

# Merge control and treatment datasets
merged_data <- merge(
  control_data,
  treatment_data,
  by = c("Ref No", "Future_Mid-point", "Climate scenario",  
         "Site(location)", "Crop","GCM/RCM","Crop Model"),
  all = FALSE
)

# Transform the data for box plot
plot_data <- merged_data %>%
  pivot_longer(
    cols = c(YieldChange_Control, YieldChange_Treatment),
    names_to = "Management_Type",
    values_to = "YieldChange"
  ) %>%
  mutate(
    Management_Group = case_when(
      Management_Type == "YieldChange_Control" ~ "No Adaptation",
      TRUE ~ `Adaptation type.y`
    )
  )

# Define **brighter & clearer** color palette
crop_colors <- c(
  "Maize" = "#377EB8",
  "Millet" = "#4DAF4A",
  "Rice" = "#FF7F00",
  "Sorghum" = "#984EA3",
  "Sunflower" = "#FFFF33",
  "Wheat" = "#E41A1C"
)

# Create **Box Plot with Crop Dots**
ggplot(plot_data, aes(
  x = YieldChange,
  y = reorder(Management_Group, YieldChange, median),
  fill = Management_Group
)) +
  geom_boxplot(
    alpha = 0.5, outlier.shape = 16, outlier.size = 1.5, 
    aes(fill = NULL), width = 0.6  # Light transparency, smaller width
  ) +
  labs(
    title = "Yield Change Distribution by Management Strategy -trt vs control",
    x = "Yield Change (%)",
    y = "Management Strategy"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 12, color = "black"),  
    axis.title.y = element_text(size = 12, face = "bold"),
    axis.title.x = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    legend.position = "none"  # Completely remove the legend
  ) 
```
